The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area

### 2023-04-28 MS_ToSubmit/MS_Burdwood_foodweb.tex
Updates Cintia Fraysse

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, large numbers (or proportions) of
endemic species, and/or high biodiversity across taxonomic levels (Hogg
et al., 2016). Indeed, the MPA N-BB was created to protect a sensitive
and biodiverse benthic habitat that was barely known (Falabella, 2017;
L. Schejter et al., 2016). The benthic community is featured by high
biomass of vulnerable and fragile species (mainly Porifera and Cnidaria)
that meet the characteristics of vulnerable marine ecosystems (Schejter
\& Albano, 2021), here defined as sites that present taxa with a minimum
biomass of 10 kg per 1200 m\textsuperscript{2} (Commission for the
Conservation of Antarctic Marine Living Resources, 2009). Also, it
provides habitat to several small-sized species (López-Gappa et al.,
2018; L. Schejter \& Bremec, 2019; Sirito, 2019), and has an important
role in the life history of fishes as a food source, refuge and nursery
area (Covatti Ale et al., 2022; Delpiani et al., 2020; Fischer et al.,
2022; García Alonso et al., 2018; Troccoli et al., 2020;
\textbf{Florencia2023?}). The maintenance of this singular community is
related to local and regional oceanographic processes, including the
circulation of the rich Malvinas (Falkland) current in the area
(Guerrero et al., 1999; A. R. Piola \& Gordon, 1989) and the upwelling
and mixing phenomena (Matano et al., 2019). The input of nutrients from
the Malvinas (Falkland) current also supports a diverse plankton
community (Guinder et al., 2020). Overall, 811 benthic and plankton
species have been identified for the MPA N-BB ecosystem, where 349 were
reported for the first time in the area in recent years (APN, 2022).
However, identifying the main species involved in the maintenance of
ecosystem services and health as well as for management and conservation
is essential.

@@

Due to a lack of trophic data resolution for some species inhabiting the
study area, we followed the concept of trophic species, here defined as
aggregated groups of taxa. In most cases, we followed it when specific
data on species, in the taxonomic sense, were not available. In some
cases, we collapsed species when taxa shared the same set of predators
and prey (trophic similarity, Martinez (1991)), one of the aggregation
methods that better preserve food web functional properties (Gauzens et
al., 2013). In addition, for endemic species (e.g.~bryozoan
\emph{Burdwoodipora paguricola}) and other species with no trophic
studies so far, we inferred their feeding interactions applying a
conservative approach that assumes that the set of prey and predators
are at some point preserved in time. In those cases we gathered
information from upper taxonomic levels (i.e.~Genus, Family, Order,
Class, Phylum) as a good proxy variable (\textbf{Morales2015?};
\textbf{Pomeranz2018?}). Details about this can be found in
Supplementary Material (Table S2).

@@

\[
CC_i = \frac{2E_i}{K_i(K_i-1)}
\] where \(E_i\) is the effective number of interactions between \(K_i\)
nearest-neighbor nodes of node \(i\) and the maximal possible number of
such interactions (Newman, 2003). To test whether the food web presented
the small-world pattern, we compared the empirical values of CPL and CC
with those resulting from 1000 randomly generated networks with the same
size (S) and number of interactions (L), following the method proposed
by Marina et al. (2018).

@@

Food web connectance is a feature that resumes the complexity of the
network, but more importantly, it is an emergent property of pairwise
species interactions (Poisot \& Gravel, 2014). It contains information
regarding how interactions within an ecological network are distributed
and predicts reasonably well key dynamical properties of ecological
networks (Dunne et al., 2002a). Complex marine food webs (i.e.~with more
than 25 trophic species) show connectance values ranging from 0.01 -
0.27 (Marina et al., 2018). In particular, food webs from high-latitude
regions tend to exhibit a connectance closer to the minimum (between
0.01 and 0.05) (Kortsch et al., 2015; Rodriguez et al., 2022; Santana et
al., 2013). Whether food webs display a low or a high connectance helps
to better comprehend ecosystem's synthetic properties like robustness.
In this sense, empirical analyses support the notion that
highly-connected ecological networks are robust against external
perturbations such as the introduction of new (e.g., invasive) species
(Smith-Ramesh et al., 2017) as well as species removal (e.g., local
extinction) (Dunne et al., 2002b; Montoya \& Solé, 2003). The
connectance of the food web of the MPA Namuncurá - Banco Burdwood (0.01)
is one of the lowest reported so far for these regions; in particular,
it appears to be five times lower than that of Beagle Channel, an
adjacent coastal area (Rodriguez et al., 2022).

@@

It is suggested that the small-world pattern, i.e., a network with short
path length and high clustering coefficient, is not frequent in complex
marine food webs, mainly due to a low clustering coefficient compared to
random networks (Dunne et al., 2002c; Marina et al., 2018). However, the
food web of the MPA N-BB does display a small-world pattern.
Consequences of this could be of great importance in recognizing
evolutionary paths and the vulnerability to perturbations (Montoya \&
Solé, 2002). On the one hand, a short path length implies a rapid spread
of an impact (e.g., contaminant, population fluctuation, local
extinction) throughout the network and, at the same time, more
potentially adaptive dynamics in the face of external perturbations
(Montoya \& Solé, 2002; Williams et al., 2002). On the other hand, a
high clustering coefficient indicates the formation of subnetworks
composed only by the neighbours of particular species. This translates
into a greater resistance of the network due to the confinement of
perturbations mainly within subnetworks and not spreading between them
(Heer et al., 2020; Kortsch et al., 2019). Overall, a small-world
topology provides ecological networks with greater resilience and
resistance (Bornatowski et al., 2017; Dormann et al., 2017).

@@

The degree distribution allows to identify important species, such as
potential keystone species (Dunne et al., 2002b; Solé \& Montoya, 2001),
generalist/specialist species, and dominant food sources (Kondoh et al.,
2010).

We have identified that the majority of the consumers in the food web of
the MPA N-BB either has a narrow diet or are specialists, while few
present a broad diet. The most evident generalist species are
\emph{Patagonotothen guntheri} (Covatti Ale et al., 2022), \emph{P.
ramsayi} (Fischer et al., 2022), \emph{Dissostichus eleginoides}
(Troccoli et al., 2020), \emph{Bathyraja brachyurops}
(\textbf{Bellegia2008?}), and \emph{B. griseocauda}
(\textbf{Bellegia2014?}), with more than 25 potential prey. Since these
species present mid-trophic positions in the food web (3-4), acting as
predator and prey, they might be important links between lower and
higher trophic levels. This is in agreement with the sole analysis,
using stable isotopes, that exists so far for the trophic structure of
the MPA N-BB (Riccialdelli et al., 2020), and resembles other
high-latitude marine systems of the Southwest Atlantic and Antarctic
regions (Arkhipkin \& Laptikhovsky, 2013; \textbf{Marina2018?}). The
importance of these particular generalist species also arises since they
feed on the benthic and pelagic habitats (Covatti Ale et al., 2022;
Fischer et al., 2022; Troccoli et al., 2020), linking these realms and
contributing to the vertical carbon flow.

On the other hand, few prey are consumed by many predators in the food
web of the MPA N-BB. This suggests that there are dominant food sources
from which most consumers depend on, and from where the ecosystem energy
is being transferred to the upper trophic levels. The most demanded
source we identified in this study (i.e.~detritus) supports the abundant
benthic community of filter-feeders (L. Schejter et al., 2016).
Furthermore, we found that the second and third-most consumed prey were
diatoms and species of \emph{Euphausia}, respectively, which are
essential sources for the diverse zooplankton community (Spinelli et
al., 2020), mid-TL consumers like the Fuegian sprat \emph{Sprattus
fuegensis} (Padovani et al., 2021) and \emph{Patagonotothen ramsayi}
(Fischer et al., 2022), and top predators (Raya Rey, 2022; Riccialdelli,
2022).

@@

Describing species' roles in food webs provides a toolbox with which to
assess the significance of species in terms of community's functioning
and overall stability (Cirtwill et al., 2018; Thébault \& Fontaine,
2010). We used a range of descriptors to characterise the dynamic and
multifaceted nature of the species forming the MPA N-BB food web.

Closeness and betweenness are defined as ``meso-scale'' properties
because they consider direct and indirect interactions, therefore
describing the focal species' ability to influence the rest of the
species of the food web (Lai et al., 2012). Closeness quantifies how
many steps away species \(i\) is from all other species in the food web,
and is proportional to how rapidly the indirect effects of the focal
species can spread to other species in the network (Scotti \& Jordán,
2010). In the food web of the MPA N-BB, low-TL consumers arise as
important in this regard: species of the zooplankton community,
\emph{Calanus} and \emph{Euphausia}, and Brachiopoda. Any perturbation
affecting these species, such as the recently confirmed contaminants
mercury (Fioramonti et al., 2022) and microplastics (Cossi et al., 2021;
Di Mauro et al., 2022) for the area, should be of concern since it might
reach many other species in the food web. Otherwise, betweenness
measures the number of shortest paths between species, providing
information on the importance of species as ``bridges'' for energy
transfer: a species with high betweenness takes part in more food chains
and therefore affects more energy flows (Scotti \& Jordán, 2010). We
have identified the longtail southern cod \emph{Patagonotothen ramsayi}
as the most important species in this sense. Moreover, in light of our
analysis, other species like the Patagonian toothfish \emph{Dissostichus
eleginoides}, the Patagonian cod \emph{Salilota australis}, the
yellowfin notothen \emph{Patagonotothen guntheri}, and the Patagonian
longfin squid \emph{Doryteuthis gahi} should be considered as relevant
in the energy transfer in the ecosystem. All these species have a
mid-trophic position in the food web, supporting the `wasp-waist'
control hypothesis for the MPA N-BB (Riccialdelli et al., 2020).

Ecosystems with a pronounced `wasp-waist' structure are suggested to
present a high trophic redundancy, since a large proportion of species
would show similar trophic habits (Cury et al., 2000). The significant
negative relationship we found between trophic similarity and trophic
level enhances the hypothesis of functional redundancy at low and mid-TL
species compared to higher TL species for the MPA N-BB food web
(Riccialdelli et al., 2020). At the same time, our results highlight the
uniqueness in terms of trophic role for high-TL predators. Here, not
only the expected pelagic animals such as marine mammals and sea birds
arise as relevant, but also benthic vertebrate (chondrichthyans
\emph{Bathyraja macloviana} and \emph{Squalus acanthias}) and
invertebrate species (echinoderms \emph{Diplopteraster clarki} and
\emph{Pteraster sp}) are noteworthy. The role that such species play in
the MPA N-BB ecosystem is unique and perturbations on them might result
in unprecedented changes at the trophic structure and functioning level.

Species' role can be also assessed in a module-based context. Among the
varying numbers of topological roles in which species can be divided,
two are remarkable: `module connector' and `network connector'. Here
again the longtail southern cod \emph{Patagonotothen ramsayi} is
emphasized as an influential species for the connectivity of the food
web, which agrees with the suggested `wasp-waist' control for the MPA
N-BB ecosystem (Riccialdelli et al., 2020). But most importantly, our
results broaden this issue by pointing out that there are other species,
belonging to a wide range of trophic positions (1 to 3.86) and
representing 20 different functional groups, that should be considered.
Thus, we propose that it is the diversity of species (benthic and
pelagic) what maintains the connectivity of the food web, therefore
contributing to the trophic structure and ecosystem's stability.

@@

Some species of sessile suspension feeders in high-latitude marine
ecosystems, such as sponges, ascidians and octocorals, avoid predation
by producing secondary metabolites that function as a chemical defense
(\textbf{NunezPons2010?}; \textbf{Moles2015?}; \textbf{Prieto2022?}).
Although this was not reported at the MPA N-BB, there are few studies
that reported it other locations in species that were found in the MPA
N-BB (\textbf{RojoDeAlmeida2010?}). Thus, the interactions that we
considered between these species and their predators should be taken
with caution.

@@

Spatial variation of study area: Banco Burdwood west-east gradient. In
the oceanic BB area, complex oceanographic conditions generate an
internal spatial heterogeneity, mainly along its longitudinal axis
(Matano et al., 2019), and consequently plankton communities respond
with different spatial distributions (Spinelli et al., 2020). Kortsch et
al. (2019)

@@

\begin{itemize}
\tightlist
\item
  Incorporate habitat heterogeneity, since benthic community in the BB
  and deep flanks is different (Schejter \& Albano, 2021). El caso de
  las ascidias (Tatián)
\end{itemize}

This might lead to distinct food web properties, hence differences in
ecosystem functioning and response to anthropogenic and environmental
perturbations. Cordone et al. (2020)

\begin{itemize}
\item
  Incorporate species traits, like body size and mass, which are known
  to be important drivers in predator-prey interactions.
\item
  Simulate perturbations on food web and target species: mercury
  transfer, microplastic pollution, fisheries. Cossi et al. (2021); Di
  Mauro et al. (2022); Fioramonti et al. (2022)
\item
  Estimation of interaction strength considering species and interaction
  traits (i.e.~body size, body mass, interaction dimensionality) to gain
  insights into food web stability and response to anthropogenic and
  environmental perturbations. Nilsson \& McCann (2016); Cirtwill et al.
  (2018); Kortsch et al. (2021)
\end{itemize}

@@

We compiled information on the species and trophic diversity of the
Marine Protected Area Namuncurá - Banco Burdwood, generating an
unprecedented, well-resolved network of trophic interactions for a
sub-Antarctic ecosystem, identifying the complexity and structure of the
system, and the main species role in a network framework. Particular
properties at the network level allowed us to identify ecosystem's
vulnerability and potential response to perturbations in the presence of
highly-connected species, with a rearrange structure maintaining their
functions due to its potential resilience and resistance.

We identified several species as important in terms of different aspects
of trophic structure and functioning, and response to perturbations
(i.e.~environmental/anthropogenic changes). On one hand, we suggest that
generalist species, mostly fishes, play a crucial role in the
ecosystem's bentho-pelagic coupling process. At the same time, we
propose that other species besides the longtail southern cod
\emph{Patagonotothen ramsayi} and the Fueguian sprat \emph{Sprattus
fuegensis} should be considered as relevant energy transfers for the
ecosystem. Finally, we argue that it is the diversity of species,
representing the benthic and pelagic habitats, what maintains the
connectivity of the food web against perturbations, therefore
contributing to the structure and stability of the ecosystem.

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, large numbers (or proportions) of
endemic species, and/or high biodiversity across taxonomic levels (Hogg
et al., 2016). Indeed, the MPA N-BB was created to protect a sensitive
and biodiverse benthic habitat that was barely known (Falabella, 2017;
L. Schejter et al., 2016). The benthic community is featured by high
biomass of vulnerable and fragile species (mainly Porifera and Cnidaria)
that meet the characteristics of vulnerable marine ecosystems (Schejter
\& Albano, 2021), here defined as sites that present taxa with a minimum
biomass of 10 kg per 1200 m\textsuperscript{2} (Commission for the
Conservation of Antarctic Marine Living Resources, 2009). Also, it
provides habitat to several small-sized species (López-Gappa et al.,
2018; L. Schejter \& Bremec, 2019; Sirito, 2019), and has an important
role in the life history of fishes as a food source, refuge and nursery
area (Covatti Ale et al., 2022; Delpiani et al., 2020; Fischer et al.,
2022; Florencia et al., 2023; García Alonso et al., 2018; Troccoli et
al., 2020). The maintenance of this singular community is related to
local and regional oceanographic processes, including the circulation of
the rich Malvinas (Falkland) current in the area (Guerrero et al., 1999;
A. R. Piola \& Gordon, 1989) and the upwelling and mixing phenomena
(Matano et al., 2019). The input of nutrients from the Malvinas
(Falkland) current also supports a diverse plankton community (Guinder
et al., 2020). Overall, 811 benthic and plankton species have been
identified for the MPA N-BB ecosystem, where 349 were reported for the
first time in the area in recent years (APN, 2022). However, identifying
the main species involved in the maintenance of ecosystem services and
health as well as for management and conservation is essential.

@@

Due to a lack of trophic data resolution for some species inhabiting the
study area, we followed the concept of trophic species, here defined as
aggregated groups of taxa. In most cases, we followed it when specific
data on species, in the taxonomic sense, were not available. In some
cases, we collapsed species when taxa shared the same set of predators
and prey (trophic similarity, Martinez (1991)), one of the aggregation
methods that better preserve food web functional properties (Gauzens et
al., 2013). In addition, for endemic species (e.g.~bryozoan
\emph{Burdwoodipora paguricola}) and other species with no trophic
studies so far, we inferred their feeding interactions applying a
conservative approach that assumes that the set of prey and predators
are at some point preserved in time. In those cases we gathered
information from upper taxonomic levels (i.e.~Genus, Family, Order,
Class, Phylum) as a good proxy variable (Morales-Castilla et al., 2015;
Pomeranz et al., 2019). Details about this can be found in Supplementary
Material (Table S2).

@@

\[
CC_i = \frac{2E_i}{K_i(K_i-1)}
\] where \(E_i\) is the effective number of interactions between \(K_i\)
nearest-neighbor nodes of node \(i\) and the maximal possible number of
such interactions (Newman, 2003). To test whether the food web presented
the small-world pattern, we compared the empirical values of CPL and CC
with those resulting from 1000 randomly generated networks with the same
size (S) and number of interactions (L), following the method proposed
by Marina, Saravia, et al. (2018).

@@

Food web connectance is a feature that resumes the complexity of the
network, but more importantly, it is an emergent property of pairwise
species interactions (Poisot \& Gravel, 2014). It contains information
regarding how interactions within an ecological network are distributed
and predicts reasonably well key dynamical properties of ecological
networks (Dunne et al., 2002a). Complex marine food webs (i.e.~with more
than 25 trophic species) show connectance values ranging from 0.01 -
0.27 (Marina, Saravia, et al., 2018). In particular, food webs from
high-latitude regions tend to exhibit a connectance closer to the
minimum (between 0.01 and 0.05) (Kortsch et al., 2015; Rodriguez et al.,
2022; Santana et al., 2013). Whether food webs display a low or a high
connectance helps to better comprehend ecosystem's synthetic properties
like robustness. In this sense, empirical analyses support the notion
that highly-connected ecological networks are robust against external
perturbations such as the introduction of new (e.g., invasive) species
(Smith-Ramesh et al., 2017) as well as species removal (e.g., local
extinction) (Dunne et al., 2002b; Montoya \& Solé, 2003). The
connectance of the food web of the MPA Namuncurá - Banco Burdwood (0.01)
is one of the lowest reported so far for these regions; in particular,
it appears to be five times lower than that of Beagle Channel, an
adjacent coastal area (Rodriguez et al., 2022).

@@

It is suggested that the small-world pattern, i.e., a network with short
path length and high clustering coefficient, is not frequent in complex
marine food webs, mainly due to a low clustering coefficient compared to
random networks (Dunne et al., 2002c; Marina, Saravia, et al., 2018).
However, the food web of the MPA N-BB does display a small-world
pattern. Consequences of this could be of great importance in
recognizing evolutionary paths and the vulnerability to perturbations
(Montoya \& Solé, 2002). On the one hand, a short path length implies a
rapid spread of an impact (e.g., contaminant, population fluctuation,
local extinction) throughout the network and, at the same time, more
potentially adaptive dynamics in the face of external perturbations
(Montoya \& Solé, 2002; Williams et al., 2002). On the other hand, a
high clustering coefficient indicates the formation of subnetworks
composed only by the neighbours of particular species. This translates
into a greater resistance of the network due to the confinement of
perturbations mainly within subnetworks and not spreading between them
(Heer et al., 2020; Kortsch et al., 2019). Overall, a small-world
topology provides ecological networks with greater resilience and
resistance (Bornatowski et al., 2017; Dormann et al., 2017).

@@

The degree distribution allows identifying important species, such as
potential keystone species (Dunne et al., 2002b; Solé \& Montoya, 2001),
generalist/specialist species, and dominant food sources (Kondoh et al.,
2010).

We have identified that most of the consumers in the food web of the MPA
N-BB either has a narrow diet or are specialists, while few present a
broad diet. The most evident generalist species are \emph{Patagonotothen
guntheri} (Covatti Ale et al., 2022), \emph{P. ramsayi} (Fischer et al.,
2022), \emph{Dissostichus eleginoides} (Troccoli et al., 2020),
\emph{Bathyraja brachyurops} (Belleggia et al., 2008), and \emph{B.
griseocauda} (Bellegia et al., 2014), with more than 25 potential prey.
Since these species present mid-trophic positions in the food web (3-4),
acting as predator and prey, they might be important links between lower
and higher trophic levels. This result is in agreement with the sole
analysis, using stable isotopes, that exists so far for the trophic
structure of the MPA N-BB (Riccialdelli et al., 2020), and resembles
other high-latitude marine systems of the Southwest Atlantic and
Antarctic regions (Arkhipkin \& Laptikhovsky, 2013; Marina, Salinas, et
al., 2018). The importance of these particular generalist species also
arises since they feed on the benthic and pelagic habitats (Covatti Ale
et al., 2022; Fischer et al., 2022; Troccoli et al., 2020), linking
these realms and contributing to the vertical carbon flow.

On the other hand, a low number of prey are consumed by many predators
in the food web of the MPA N-BB. This suggests that there are dominant
food sources from which most consumers depend on and from where the
ecosystem energy is being transferred to the upper trophic levels. The
most demanded source we identified in this study (i.e.~detritus)
supports the abundant benthic community of filter-feeders (L. Schejter
et al., 2016). Furthermore, we found that the second and third-most
consumed prey were diatoms and species of \emph{Euphausia},
respectively, which are essential sources for the diverse zooplankton
community (Spinelli et al., 2020), mid-TL consumers like the Fuegian
sprat \emph{Sprattus fuegensis} (Padovani et al., 2021) and
\emph{Patagonotothen ramsayi} (Fischer et al., 2022), and top predators
(Raya Rey, 2022; Riccialdelli, 2022).

@@

Describing species' roles in food webs provides a toolbox to assess the
significance of species in terms of community's functioning and overall
stability (Cirtwill et al., 2018; Thébault \& Fontaine, 2010). We used a
range of descriptors to characterise the dynamic and multifaceted nature
of the species forming the MPA N-BB food web.

Closeness and betweenness are defined as ``mesoscale'' properties
because they consider direct and indirect interactions, therefore
describing the focal species' ability to influence the rest of the
species of the food web (Lai et al., 2012). Closeness quantifies how
many steps away species \(i\) is from all other species in the food web,
and is proportional to how rapidly the indirect effects of the focal
species can spread to other species in the network (Scotti \& Jordán,
2010). In the food web of the MPA N-BB, low-TL consumers arise as
important in this regard: species of the zooplankton community,
\emph{Calanus} and \emph{Euphausia}, and Brachiopoda. Any perturbation
affecting these species, such as the recently confirmed contaminants
mercury (Fioramonti et al., 2022) and microplastics (Cossi et al., 2021;
Di Mauro et al., 2022) for the area, should be of concern since it might
reach many other species in the food web. Otherwise, betweenness
measures the number of shortest paths between species, providing
information on the importance of species as ``bridges'' for energy
transfer: a species with high betweenness takes part in more food chains
and therefore affects more energy flows (Scotti \& Jordán, 2010). We
have identified the longtail southern cod \emph{Patagonotothen ramsayi}
as the most important species in this sense. Moreover, in light of our
analysis, other species like the Patagonian toothfish \emph{Dissostichus
eleginoides}, the Patagonian cod \emph{Salilota australis}, the
yellowfin notothen \emph{Patagonotothen guntheri}, and the Patagonian
longfin squid \emph{Doryteuthis gahi} should be considered as relevant
in the energy transfer in the ecosystem. All these species have a
mid-trophic position in the food web, supporting the `wasp-waist'
control hypothesis for the MPA N-BB (Riccialdelli et al., 2020).

Ecosystems with a pronounced `wasp-waist' structure are suggested to
present a high trophic redundancy, since many species would show similar
trophic habits (Cury et al., 2000). The significant negative
relationship between trophic similarity and trophic level enhances the
hypothesis of functional redundancy at low and mid-TL species compared
to higher TL species for the MPA N-BB food web (Riccialdelli et al.,
2020). At the same time, our results highlight the uniqueness in terms
of the trophic role of high-TL predators. Here, not only the expected
pelagic animals such as marine mammals and sea birds arise as relevant,
but also benthic vertebrate (chondrichthyans \emph{Bathyraja macloviana}
and \emph{Squalus acanthias}) and invertebrate species (echinoderms
\emph{Diplopteraster clarki} and \emph{Pteraster sp}) are noteworthy.
The role that such species play in the MPA N-BB ecosystem is unique and
perturbations on them might result in unprecedented changes at the
trophic structure and functioning level.

Species' role can also be assessed in a module-based context. Among the
varying numbers of topological roles in which species can be divided,
two are remarkable: `module connector' and `network connector'. Here
again, the longtail southern cod \emph{Patagonotothen ramsayi} is
emphasized as an influential species for the connectivity of the food
web, which agrees with the suggested `wasp-waist' control for the MPA
N-BB ecosystem (Riccialdelli et al., 2020). But most importantly, our
results broaden this issue by pointing out that there are other species,
belonging to a wide range of trophic positions (1 to 3.86) and
representing 20 different functional groups, that should be considered.
Thus, we propose that the diversity of species (benthic and pelagic)
maintains the connectivity of the food web, therefore contributing to
the trophic structure and ecosystem's stability.

@@

Some species of sessile suspension feeders in high-latitude marine
ecosystems, such as sponges, ascidians and octocorals, avoid predation
by producing secondary metabolites that function as a chemical defense
(Moles et al., 2015; Núñez-Pons et al., 2010; Prieto et al., 2022).
Although this was not reported at the MPA N-BB, there are few studies
that reported it other locations in species that were found in the MPA
N-BB (Rojo de Almeida et al., 2010). Thus, the interactions that we
considered between these species and their predators should be taken
with caution.

@@

Spatial variation of study area: Banco Burdwood west-east gradient. In
the oceanic BB area, complex oceanographic conditions generate an
internal spatial heterogeneity, mainly along its longitudinal axis
(Matano et al., 2019), and consequently, plankton communities respond
with different spatial distributions (Spinelli et al., 2020). Kortsch et
al. (2019)

@@

\begin{itemize}
\item
  Incorporate habitat heterogeneity, since the benthic community in the
  BB and deep flanks differs (Schejter \& Albano, 2021). This
  heterogeneity might lead to distinct food web properties, hence
  differences in ecosystem functioning and response to anthropogenic and
  environmental perturbations (Cordone et al., 2020). El caso de las
  ascidias (Tatián)
\item
  Incorporate species traits, like body size and mass, which are known
  to be important drivers in predator-prey interactions.
\item
  Simulate perturbations on food web and target species: mercury
  transfer, microplastic pollution, fisheries. Cossi et al. (2021); Di
  Mauro et al. (2022); Fioramonti et al. (2022)
\item
  Estimation of interaction strength considering species and interaction
  traits (i.e.~body size, body mass, interaction dimensionality) to gain
  insights into food web stability and response to anthropogenic and
  environmental perturbations. Nilsson \& McCann (2016); Cirtwill et al.
  (2018); Kortsch et al. (2021)
\end{itemize}

@@

We compiled information on the species and trophic diversity of the
Marine Protected Area Namuncurá - Banco Burdwood, generating an
unprecedented, well-resolved network of trophic interactions for a
sub-Antarctic ecosystem, identifying the complexity and structure of the
system, and the main species role in a network framework. Particular
properties at the network level allowed us to identify the ecosystem's
vulnerability and potential response to perturbations in the presence of
highly-connected species, with a rearranged structure maintaining their
functions due to its potential resilience and resistance.

We identified several species as important regarding different aspects
of trophic structure and functioning, and response to perturbations
(i.e.~environmental/anthropogenic changes). On the one hand, we suggest
that generalist species, mainly fishes, play a crucial role in the
ecosystem's bentho-pelagic coupling process. At the same time, we
propose that other species besides the longtail southern cod
\emph{Patagonotothen ramsayi} and the Fueguian sprat \emph{Sprattus
fuegensis} should be considered relevant energy transfers for the
ecosystem. Finally, we argue that it is the diversity of species,
representing the benthic and pelagic habitats, that maintains the
connectivity of the food web against perturbations, therefore
contributing to the structure and stability of the ecosystem.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Belleggia2019}{}}%
Belleggia, M., Battagliotti, C., Cortés, F., \& Colonello, J. H. (2019).
Feeding together: A global diet analysis of twenty-three species of
chondrichthyes on a feeding ground area. \emph{Hydrobiologia},
\emph{842}(1), 77--99. \url{https://doi.org/10.1007/s10750-019-04027-w}

\leavevmode\vadjust pre{\hypertarget{ref-Belleggia2008a}{}}%
Belleggia, M., Mabragaña, E., Figueroa, D. E., Scenna, L. B., Barbini,
S. A., \& Astarloa, J. M. D. de. (2008). Food habits of the broad nose
skate, {\emph{Bathyraja}}{ \emph{Brachyurops}} ({Chondrichthyes},
{Rajidae}), in the south-west {Atlantic}. \emph{Scientia Marina},
\emph{72}(4), 701--710.
\url{https://doi.org/10.3989/scimar.2008.72n4701}

\leavevmode\vadjust pre{\hypertarget{ref-Bellegia2014}{}}%
Bellegia, M., Scenna, L., Barbini, S. A., Figueroa, D. E., \& Díaz de
Astarloa, J. M. (2014). \emph{The diets of four {Bathyraja} skates
({Elasmobranchii}, {Arhynchobatidae}) from the {Southwest Atlantic}}.
\url{https://doi.org/10.26028/CYBIUM/2014-384-012}

@@

\leavevmode\vadjust pre{\hypertarget{ref-AlvarezColombo2018}{}}%
Colombo, G. Álvarez. (2018). \emph{{Informe de Campaña Namuncurá-Banco
Burdwood: \textquestiondown Area de cría de peces australes?}} (p. 83)
{[}\{Field campaign Report\}{]}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Dellabianca2015}{}}%
Dellabianca, N. A. (2015). \emph{{Informe de Campaña {``Namuncurá-Banco
Burdwood''}}} (p. 48) {[}\{Field campaign Report\}{]}.

\leavevmode\vadjust pre{\hypertarget{ref-Dellabianca2018}{}}%
Dellabianca, N. A. (2018). \emph{{Informe de Campaña AMP Namuncurá
\textendash{} Banco Burdwood: Comprendiendo la bomba biológica de
carbono}} (p. 101) {[}\{Field campaign Report\}{]}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Florencia2023}{}}%
Florencia, M., Vazquez, D. M., Gabbanelli, V., Díaz de Astarloa, J. M.,
\& Mabragaña, E. (2023). Chondrichthyans from the southern tip of {South
America} with emphasis on the marine protected area {Namuncurá-Burdwood
Bank}: Exploring egg nursery grounds. \emph{Polar Biology}.
\url{https://doi.org/10.1007/s00300-023-03128-6}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Giberto2017}{}}%
Giberto, D. A., Belleggia, M., \& Bremec, C. S. (2017). \emph{El bentos
como alimento de peces comerciales.} {Instituto Nacional de
Investigación y Desarrollo Pesquero (INIDEP)}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Kreps2016}{}}%
Kreps, G. (2016). \emph{{Informe de Campaña AMP Namuncurá - Banco
Burdwood: Primavera 2016}} (p. 103) {[}\{Field campaign Report\}{]}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Marina2018}{}}%
Marina, T. I., Salinas, V., Cordone, G., Campana, G., Moreira, E.,
Deregibus, D., Torre, L., Sahade, R., Tatián, M., Barrera Oro, E., De
Troch, M., Doyle, S., Quartino, M. L., Saravia, L. A., \& Momo, F. R.
(2018). The {Food Web} of {Potter Cove} ({Antarctica}): Complexity,
structure and function. \emph{Estuarine, Coastal and Shelf Science},
\emph{200}, 141--151. \url{https://doi.org/10.1016/j.ecss.2017.10.015}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Martin2015}{}}%
Martin, J. (2015). \emph{{Informe de Campaña Banco Burdwood/Namuncurá
\textendash{} Tierra del Fuego}} (p. 78) {[}\{Field campaign
Report\}{]}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Moles2015}{}}%
Moles, J., Núñez-Pons, L., Taboada, S., Figuerola, B., Cristobo, J., \&
Avila, C. (2015). Anti-predatory chemical defences in {Antarctic}
benthic fauna. \emph{Marine Biology}, \emph{162}(9), 1813--1821.
\url{https://doi.org/10.1007/s00227-015-2714-9}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Morales-Castilla2015}{}}%
Morales-Castilla, I., Matias, M. G., Gravel, D., \& Araújo, M. B.
(2015). Inferring biotic interactions from proxies. \emph{Trends in
Ecology \& Evolution}, \emph{30}(6), 347--356.
\url{https://doi.org/10.1016/j.tree.2015.03.014}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Nunez-Pons2010}{}}%
Núñez-Pons, L., Forestieri, R., Nieto, R. M., Varela, M., Nappo, M.,
Rodríguez, J., Jiménez, C., Castelluccio, F., Carbone, M., Ramos-Espla,
A., Gavagnin, M., \& Avila, C. (2010). Chemical defenses of tunicates of
the genus {Aplidium} from the {Weddell Sea} ({Antarctica}). \emph{Polar
Biology}, \emph{33}(10), 1319--1329.
\url{https://doi.org/10.1007/s00300-010-0819-7}

\leavevmode\vadjust pre{\hypertarget{ref-Ospina-Alvarez2020}{}}%
Ospina-Alvarez, A., de Juan, S., Alós, J., Basterretxea, G.,
Alonso-Fernández, A., Follana-Berná, G., Palmer, M., \& Catalán, I.
(2020). {MPA} network design based on graph theory and emergent
properties of larval dispersal. \emph{Marine Ecology Progress Series},
\emph{650}, 309--326. \url{https://doi.org/10.3354/meps13399}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Pomeranz2019}{}}%
Pomeranz, J. P. F., Thompson, R. M., Poisot, T., \& Harding, J. S.
(2019). Inferring predator\textendash prey interactions in food webs.
\emph{Methods in Ecology and Evolution}, \emph{10}(3), 356--367.
\url{https://doi.org/10.1111/2041-210X.13125}

\leavevmode\vadjust pre{\hypertarget{ref-Prieto2022}{}}%
Prieto, I. M., Paola, A., Pérez, M., García, M., Blustein, G., Schejter,
L., \& Palermo, J. A. (2022). Antifouling {Diterpenoids} from the
{Sponge} {\emph{Dendrilla}}{ \emph{Antarctica}}. \emph{Chemistry \&
Biodiversity}, \emph{19}(2).
\url{https://doi.org/10.1002/cbdv.202100618}

@@

\leavevmode\vadjust pre{\hypertarget{ref-RojodeAlmeida2010}{}}%
Rojo de Almeida, M. T., Siless, G. E., Perez, C. D., Veloso, M. J.,
Schejter, L., Puricelli, L., \& Palermo, J. A. (2010). Dolabellane
{Diterpenoids} from the {South Atlantic Gorgonian} {\emph{Convexella}}{
\emph{Magelhaenica}}. \emph{Journal of Natural Products}, \emph{73}(10),
1714--1717. \url{https://doi.org/10.1021/np100337j}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Schejter2017}{}}%
Schejter, L. (2017). \emph{{Informe de Campaña Banco Burdwood}} (p. 296)
{[}\{Field campaign Report\}{]}.
